---
title: "Interactive Maps with Leaflet"
subtitle: "Using Leaflet: A Tutorial"
author: "@elenamurray, @francobastida"
date: "`r format(Sys.time(), '%B %d, %Y | %H:%M:%S | %Z')`"
output:
  html_document:
    toc: yes
    df_print: paged
    number_sections: no
    highlight: tango
    theme: sandstone
    includes:
      after_body: null
    toc_depth: 3
    toc_float: yes
    css: custom.css
    self_contained: no
  word_document:
    toc: yes
    toc_depth: '3'
institute: Hertie School
---

```{=html}
<style>
div.answer {background-color:#f3f0ff; border-radius: 5px; padding: 5px;}
</style>
```
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

<br>

The contents of this session will introduce `leaflet`  as a data visualization tool that can help you identify patterns, trends, and relationships that may not be obvious in raw data. This can add value to your toolbox as a data scientist or practitioner supporting policymakers and other stakeholders.

**What are the objectives of this learning session?**

The goals of this session are to:

1. equip you with conceptual knowledge to use the Leaflet package and integrate your own data sets;
2. show you some key functions of the package to add markers, polygons, and other interesting layers; and
3. provide you with concrete examples through a tutorial as well as additional resources.

# Getting Started with a Base Map

First, install the packages that we need to use.

```{r, include = T}
#Load the necessary packages

library(htmlwidgets)
library(leaflet)
library(tidyverse)
library(sf)
library(jsonlite)
library(RColorBrewer)
```

## Setting up a Base Map

Let's start by adding the base tiles of the map, and adding one marker for a point of interest. Remember:

-   `addTiles()` add your base map layer using OpenStreetMap

-   `addMarkers()` pinpoint a point of interest on the map

What's a place that you have top of mind? You can find the coordinates (the longitude and the latitude) to many points of interest on the [Open Street Map Website](https://www.openstreetmap.org/). 

```{r}
#Adding a base map and coordinates
example <- leaflet() %>%
  addTiles() %>%  #Add default OpenStreetMap map tiles
  addMarkers(lng=4.3007, lat=52.0705,  #Include desired coordinates 
  					 popup="Hello the Hague!") #Include pop-up message 

example  #Print the map
```
<br>

# Changing the Base Map 

Did you know that OpenStreetMap is not the only base map that you can use? Many other free, third-party base maps can be added using the `addProviderTiles()` function, using the leaflet-providers plugin. You can find the full list of options [here](https://leaflet-extras.github.io/leaflet-providers/preview/index.html). 

**Leaflet** also provides a list of all the third-party tile providers that are supported by the plugin. This enables you to use auto-completion feature of your favorite IDE (like RStudio) and not have to remember or look up tile providers; just type `providers$` and choose from one of the options. You can also use `names(providers)` to view all of the options.

```{r}
#Using different map options e.g. OpenRailway
example %>% addProviderTiles(providers$OpenRailwayMap)
```

## Exercise 1

For this exercise, you will warm up and use the functions to create a map that:

-  looks up a unique location
-  adds a marker with a message
-  includes a base map from the providers' list

```{r}
#Please show your Map 1 code here
  
```

<br>

# Importing Data 

We will now import the data sets we'll be using for the next exercise. We are using a GeoJSON file containing the location of all the **Christmas Markets in Berlin**, which was taken from the [Berlin Open Data Portal](https://daten.berlin.de/datensaetze/simple_search_wwwberlindesenwebservicemaerktefesteweihnachtsmaerkte).

We will also be using Polygon data showing the different neighbourhoods of Berlin, in this case we will use GeoJSON data from [Geoportal Berlin (2017)](https://daten.odis-berlin.de/de/dataset/bezirksgrenzen/).

```{r}
#Load the GeoJSON data file for the Christmas Markets
markets_data <- st_read("Leaflet_Tutorial_files/data/markets.geojson")

#Load the Polygons data file
polygon_data <- read_sf("leaflet_workshopslides_files/data/bezirksgrenzen.geojson")

```

<br>

### Adding Markers

Let's use the data set to add a marker for each Christmas market on a map of Berlin. Within this data set, there is columnn titled 'Data' that contains a JSON string with details about each Christmas market, including the name, address, and other attributes. 

To extract and display the relevant information from the data column, we will need to first parse the JSON strings within that column, and then we can create the Leaflet map. 

```{r}
#Parse the JSON data in the 'data' column
markets_data$data_parsed <- lapply(markets_data$data, fromJSON)

#Convert the list into a data frame and bind it to the original dataset
markets_data_parsed <- do.call(rbind, lapply(markets_data$data_parsed, as.data.frame))

#Combine the parsed data with the original dataset
markets_full <- cbind(markets_data, markets_data_parsed)

#Create the leaflet map
leaflet(markets_full) %>%
  addTiles() %>%
  setView(lng = 13.4050, lat = 52.52, zoom = 12) %>%
  addMarkers(
    data = markets_full, 
    lng = st_coordinates(markets_full)[,1],   # Longitude
    lat = st_coordinates(markets_full)[,2],    # Latitude
    popup = ~paste("<strong>Market:</strong> ", markets_full$name, 
                   "<br><strong>Address:</strong> ", markets_full$strasse, 
                   "<br><strong>Opening hours:</strong> ", markets_full$oeffnungszeiten,
                   "<br><strong>Organizer:</strong> ", markets_full$veranstalter,
                   "<br><strong>Email:</strong> ", markets_full$email)
  )
```

<br>

## Adding Polygons

We can now add the polygon data to this map, so we can easily see which markets are in which neighbourhood in Berlin. 

```{r}
#Combine the two data sets
combined_data <- st_join(markets_full, polygon_data, join = st_within)

#Create Leaflet map
leaflet() %>%
  addTiles() %>%
  setView(lng = 13.4050, lat = 52.52, zoom = 12) %>%
  
#Add neighborhood polygons
  addPolygons(data = polygon_data, 
               fillColor = "lightblue", 
               color = "black", 
               weight = 1, 
               opacity = 0.5, 
               fillOpacity = 0.5, 
               highlightOptions = highlightOptions(
                 weight = 2,
                 color = "white",
                 fillOpacity = 0.7,
                 bringToFront = TRUE
               ),
               label = ~Gemeinde_name  
  ) %>%  
  
#Add market markers
  addMarkers(
    data = combined_data, 
    lng = st_coordinates(combined_data)[, 1],   # Longitude
    lat = st_coordinates(combined_data)[, 2],    # Latitude
    popup = ~paste("<strong>Market:</strong> ", combined_data$name, 
                   "<br><strong>Address:</strong> ", combined_data$strasse, 
                   "<br><strong>Opening hours:</strong> ", combined_data$oeffnungszeiten,
                   "<br><strong>Organizer:</strong> ", combined_data$veranstalter,
                   "<br><strong>Email:</strong> ", combined_data$email)
  )  
```

<br>

## Exercise 2

For this exercise, you will create a map of Berlin with the open data from [Kristof and Lukasz (2021)](https://zenodo.org/records/4446043 ) on rental listings in European cities. His example uses the `berlin_weekends.csv` file on Airbnb listings from the repo. Here, you will construct either...

- option a) map that clusters markers according to cleanliness (hint: you will have to mutate or analyze the current data frame)
- option b) map that creates a range for the level of guest satisfaction to color each district (hint: you will have to join the data)

```{r}
#Please show your Map 2 code here
  
```

<br>

There are more resources available in the [repository page](https://github.com/intro-to-data-science-24-workshop/09-leafletgraphs-murray-bastida) that we have created. Remember to check out the presentation and materials, as they could provide valuable insights for troubleshooting, examples, and inspiration. 

Happy mapping!

### Thank you!